---
title: "Summary of TMA36 cohort: clinical characteristics"
author: "Mafe Senosain"
date: "`r Sys.Date()`"
output:
    html_document:
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(RColorBrewer)
library(gridExtra)
library(reshape2)
library(jcolors)
library(ComplexHeatmap)
library(dplyr)
library(tidyr)
library(ggrepel)
library(forcats)
library(ggsignif)
library(ggpubr)
knitr::opts_chunk$set(echo = FALSE)
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/R/20_ClustAnnot_functions.R")
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/R/30_DA_functions.R")
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/R/40_DE_functions.R")
```

```{r, echo=FALSE}
# read edited CSV file as CDE_TMA36
CDE <- read.csv(file = '/Users/senosam/Documents/Massion_lab/CyTOF_summary/CDE_TMA36_2021JAN13_SA_MF.csv')
```

```{r, echo=FALSE}
pie_chart <- function(dt, col_name, plot_title){
    x <- data.frame(table(dt[col_name]))
    colnames(x) <- c('Group', 'value')
    x <- x %>% 
        arrange(desc(value)) #%>%
        #mutate(prop = label_percent(value / sum(value)))

    print(ggplot(x, aes(x = "", y = value, fill = fct_inorder(Group))) +
        geom_bar(width = 1, stat = "identity") +
        coord_polar("y", start = 0) +
        geom_label_repel(aes(label = value), size=5, show.legend = F, nudge_x = 1) +
        #geom_label_repel(aes(label = prop), size=5, show.legend = F, nudge_x = 1) + #%
        guides(fill = guide_legend(title = "Group")) +
        labs(title=plot_title) +
        theme(plot.title = element_text(hjust = 0.5, size=22)))
}

```

# Clinical characteristics distributions on TMA36 cohort
```{r, echo=FALSE}
hist(CDE$SILA, xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE, 'CANARY', 'CANARY')
pie_chart(CDE, 'Stages_simplified', 'Stage')
pie_chart(CDE, 'Hist_predominant', 'Histology')
pie_chart(CDE, 'Death_st', 'Death Status')
pie_chart(CDE, 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE, 'Progression_st', 'Progression Status')
pie_chart(CDE, 'Smoking_Status', 'Smoking Status')
```

# Clinical characteristics distributions by CANARY

## Good prognosis (Indolent)
```{r, echo=FALSE}
k <- which(CDE$CANARY == 'G')
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')
pie_chart(CDE[k,], 'Smoking_Status', 'Smoking Status')

```

## Intermediate prognosis
```{r, echo=FALSE}
k <- which(CDE$CANARY == 'I')
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')
pie_chart(CDE[k,], 'Smoking_Status', 'Smoking Status')
```

## Poor prognosis (Aggressive)
```{r, echo=FALSE}
k <- which(CDE$CANARY == 'P')
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')
pie_chart(CDE[k,], 'Smoking_Status', 'Smoking Status')
```

# Clinical characteristics distributions by SILA Score

```{r, echo=FALSE}
#http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/76-add-p-values-and-significance-levels-to-ggplots/
bp_f <- function(dt, chr_colnm, sila_colnm='SILA', main_title, x_title) {
  dt <- dt[,c(sila_colnm,chr_colnm)]
  dt <- melt(dt)
  colnames(dt)[1]<- 'condition'
  
  x <- compare_means(value ~ condition, dt, method = "wilcox.test", paired = FALSE,
                     group.by = NULL, ref.group = NULL)
  idx <- which(x$p.signif != 'ns')
  ls <- list()
  n=1
  for (i in idx) {
    ls[[n]] <- c(x$group1[idx[n]], x$group2[idx[n]])
    n=n+1
  }
  #print(data.frame(x))
  
  p <- ggplot(dt, aes(x=condition, y=value, fill=condition)) +
    geom_boxplot() +
    labs(title=main_title,x=x_title, y = "SILA score") +
    theme(plot.title = element_text(hjust = 0.5, size=20)) +
    ggsignif::geom_signif(comparisons = ls, 
       map_signif_level=TRUE)
    geom_signif(comparisons = ls, 
              map_signif_level=TRUE)
  return(p)
}

bp_f(CDE, chr_colnm='CANARY', sila_colnm='SILA', main_title='SILA score by CANARY', x_title='CANARY')
bp_f(CDE, chr_colnm='Stages_simplified', sila_colnm='SILA', main_title='SILA score by Stage', x_title='Tumor Stage')
bp_f(CDE, chr_colnm='Hist_predominant', sila_colnm='SILA', main_title='SILA score by Histology', x_title='Predominant Histology')
bp_f(CDE, chr_colnm='Death_st', sila_colnm='SILA', main_title='SILA score by Death Status', x_title='Death Status')
bp_f(CDE, chr_colnm='Recurrence_st', sila_colnm='SILA', main_title='SILA score by Recurrence Status', x_title='Recurrence Status')
bp_f(CDE, chr_colnm='Progression_st', sila_colnm='SILA', main_title='SILA score by Progression Status', x_title='Progression Status')
bp_f(CDE, chr_colnm='Smoking_Status', sila_colnm='SILA', main_title='SILA score by Smoking Status', x_title='Smoking Status')


```

```{r}
# ggpubr::ggscatter(CDE, x = 'SILA', y = c("Age_at_collection"),
#           add = "reg.line", conf.int = TRUE, combine = TRUE,
#           cor.coef = TRUE, cor.method = "spearman", add.params = list(color = 'grey'),
#           xlab = "SILA score", ylab = 'condition')
```


# New classification 

## Identifying outliers
```{r, echo=FALSE}
flag = c()
for (i in 1:nrow(CDE)){
    # if SILA >=0.5 stage shouldn't be 0, histology shouldn't be lepidic
    if (CDE$SILA[i]>=0.5) {
        if( CDE$Stages_simplified[i] %in% c('Stage 0') || 
            CDE$Hist_predominant[i] %in% c('lepidic')) {
            #cat('flag!: ', CDE$pt_ID[i], '\n')
            flag = c(flag, i)
        }
    }
    # if SILA <0.5 stage shouldn't be 3 or 4, histology shouldn't be solid or micropapillary, 
    # patient shouldn't be dead, have a recurrence or a progression, CANARY shouldn't be Poor
    if (CDE$SILA[i]<0.5) {
        if ( CDE$Stages_simplified[i] %in% c('Stage III', 'Stage IV') || 
            CDE$Hist_predominant[i] %in% c('solid', 'micropapillary') || 
            CDE$Death_st[i] == 'Yes' || 
            CDE$Recurrence_st[i] == 'Yes' || 
            CDE$Progression_st[i] == 'Yes' ||
            CDE$CANARY[i] == 'P' ){
            #cat('flag!: ', CDE$pt_ID[i], '\n')
            flag = c(flag, i)
        }
    }
}

fl_df <- CDE[flag, c('pt_ID','SILA', 'CANARY', 'Stages_simplified', 'X8th_ed_path_stage', 
    'Hist_predominant', 'Hist_other_patterns', 'Death_st', 'Recurrence_st', 
    'Progression_st')]

fl_g <- fl_df[which(fl_df$CANARY == 'G'),]
fl_i <- fl_df[which(fl_df$CANARY == 'I'),]
fl_p <- fl_df[which(fl_df$CANARY == 'P'),]
fl_df <- rbind(fl_g, fl_i, fl_p)
knitr::kable(fl_df)
```


## Option 1: Indolent < 0.5 | Aggressive  >= 0.5

```{r, echo=FALSE}
ind <- CDE[which(CDE$SILA <= 0.5),]
agg <- CDE[which(CDE$SILA > 0.5),]
n_op1 <- c(rep(0, nrow(CDE)))
```

### Indolent as SILA < 0.5
```{r, echo=FALSE}
k <- which(CDE$SILA < 0.5)
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')

i_df <- CDE[k, c('pt_ID','SILA', 'CANARY', 'Stages_simplified', 'X8th_ed_path_stage', 
    'Hist_predominant', 'Hist_other_patterns', 'Death_st', 'Recurrence_st', 
    'Progression_st')]
knitr::kable(i_df)
n_op1[k] <- 'ind'
```

### Aggressive as SILA >= 0.5
```{r, echo=FALSE}
k <- which(CDE$SILA >= 0.5)
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')

a_df <- CDE[k, c('pt_ID','SILA', 'CANARY', 'Stages_simplified', 'X8th_ed_path_stage', 
    'Hist_predominant', 'Hist_other_patterns', 'Death_st', 'Recurrence_st', 
    'Progression_st')]
knitr::kable(a_df)
n_op1[k] <- 'agg'
```

## Option 2: Indolent < 0.4 | intermediate > 0.4 <= 0.6 | Aggressive  > 0.6

```{r, echo=FALSE}
ind <- CDE[which(CDE$SILA <= 0.4),]
int <- CDE[which(CDE$SILA > 0.4 & CDE$SILA <= 0.6),]
agg <- CDE[which(CDE$SILA < 0.6),]
n_op2 <- c(rep(0, nrow(CDE)))
```

### Indolent as SILA <= 0.4
```{r, echo=FALSE}
k <- which(CDE$SILA <= 0.4)
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')
i_df <- CDE[k, c('pt_ID','SILA', 'CANARY', 'Stages_simplified', 'X8th_ed_path_stage', 
    'Hist_predominant', 'Hist_other_patterns', 'Death_st', 'Recurrence_st', 
    'Progression_st')]
knitr::kable(i_df)
n_op2[k] <- 'ind'
```

### Intermediate as SILA > 0.4 <= 0.6
```{r, echo=FALSE}
k <- which(CDE$SILA > 0.4 & CDE$SILA <= 0.6)
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')
int_df <- CDE[k, c('pt_ID','SILA', 'CANARY', 'Stages_simplified', 'X8th_ed_path_stage', 
    'Hist_predominant', 'Hist_other_patterns', 'Death_st', 'Recurrence_st', 
    'Progression_st')]
knitr::kable(int_df)
n_op2[k] <- 'int'
```

### AGGRESSIVE as SILA > 0.6
```{r, echo=FALSE}
k <- which(CDE$SILA > 0.6)
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')
a_df <- CDE[k, c('pt_ID','SILA', 'CANARY', 'Stages_simplified', 'X8th_ed_path_stage', 
    'Hist_predominant', 'Hist_other_patterns', 'Death_st', 'Recurrence_st', 
    'Progression_st')]
knitr::kable(a_df)
n_op2[k] <- 'agg'
```

```{r, echo=FALSE}
CDE['n_op1'] <- n_op1
CDE['n_op2'] <- n_op2
write.csv(CDE, "/Users/senosam/Documents/Massion_lab/CyTOF_summary/CDE_TMA36_2021JAN13_SA_MF.csv", row.names = FALSE)
```

```{r, echo=FALSE}
bp_f(CDE, chr_colnm='n_op1', sila_colnm='SILA', main_title='SILA score by Option 1', x_title='new classification')
bp_f(CDE, chr_colnm='n_op2', sila_colnm='SILA', main_title='SILA score by Option 2', x_title='new classification')
```

# Summary Omics datasets heatmap

```{r}
# Load data
df_cytof <- read.csv(file = '/Users/senosam/Documents/Massion_lab/data_integration/models/data/cytof_medianprot.csv')
df_rna <- read.csv(file = '/Users/senosam/Documents/Massion_lab/data_integration/models/data/rna_viper25.csv')
df_wes <- read.csv(file = '/Users/senosam/Documents/Massion_lab/data_integration/models/data/wes_binary.csv')
df_rad <- read.csv(file = '/Users/senosam/Documents/Massion_lab/data_integration/models/data/rad_healthmyne.csv')
```

```{r}
CDE_omics <- CDE
CDE_omics['cytof'] <- 0
CDE_omics[which(CDE$pt_ID %in% df_cytof$X),'cytof'] <- 1
CDE_omics['rnaseq'] <- 0
CDE_omics[which(CDE$pt_ID %in% df_rna$X),'rnaseq'] <- 1
CDE_omics['wes'] <- 0
CDE_omics[which(CDE$pt_ID %in% df_wes$X),'wes'] <- 1
# CDE_omics['HM'] <- 0
# CDE_omics[which(CDE$pt_ID %in% df_rad$X),'HM'] <- 1
CDE_omics['mxif'] <- 1
sc_rna <- c('12929', '13774', '7984', '14965', '12889', '8356', 
            '11918', '11522', '15002', '15467', '13634', '14428',
            '11817', '14958', '12935', '13636')
CDE_omics['scrnaseq'] <- 0
CDE_omics[which(CDE$pt_ID %in% sc_rna),'scrnaseq'] <- 1
rownames(CDE_omics) <- CDE_omics$pt_ID
CDE_omics <- CDE_omics[,86:ncol(CDE_omics)]
```

```{r}
mat = t(as.matrix(CDE_omics))

ha = HeatmapAnnotation(
  Prognosis = CDE$n_op2,
  SILA = CDE$SILA,
  Stage = CDE$Stages_simplified,
  Histology = CDE$Hist_predominant,
  Smoking_Status = CDE$Smoking_Status,
  Nodule_size_cm = CDE$Path_Nodule_Size_cm,
  Sex = CDE$Gender,
  Age = CDE$Age_at_collection,
  gp = gpar(col = "black"),
  simple_anno_size = unit(0.3, "cm"),
  annotation_name_gp = gpar(fontsize = 8) #size of annotation name
)

bp_top = HeatmapAnnotation(sum = anno_barplot(apply(CDE_omics, 1, sum), border=FALSE),
                           show_annotation_name = FALSE,
                           gp = gpar(fill =2:3, col =2:3))

ht_opt(heatmap_column_names_gp = gpar(fontsize = 5), 
    heatmap_column_title_gp = gpar(fontsize = 10, fontface = "bold"),
    heatmap_row_names_gp = gpar(fontsize = 10, fontface = "bold"),
    legend_title_gp = gpar(fontsize = 8, fontface = "bold"),
    legend_labels_gp = gpar(fontsize = 8)
)
#ht_opt(RESET = TRUE)

ht = Heatmap(mat, name = "Data \ncollected", col = c('grey', 'red'),
        column_split = factor(CDE$n_op2, levels= c('ind', 'int', 'agg')),
        column_title = c('Indolent', 'Intermediate', 'Aggressive'),
        row_labels = c('CyTOF \n(n=71)', 'RNA-Seq \n(n=78)', 'WES \n(n=57) ', 'MxIF \n(n=93)', 'scRNA-Seq \n(n=16)'),
        width = unit(18, "cm"), 
        height = unit(5, "cm"),
        #width = unit(10, "cm"), 
        #height = unit(5, "cm"),
        bottom_annotation = ha, 
        top_annotation = bp_top,
        show_row_dend = FALSE,
        show_column_dend = FALSE,
        column_order = order(CDE$SILA),
        row_order = c(1:5),
        rect_gp = gpar(col = "white", lwd = 1)
        )
draw(ht, padding = unit(c(2, 20, 2, 2), "mm"))
#draw(ht, padding = unit(c(6, 20, 6, 20), "cm"))
```

```{r}
k = which(CDE$pt_ID %in% sc_rna)
CDE_scRNA = CDE[k,]
di_ii <- grep('Date', colnames(CDE_scRNA))
CDE_scRNA[di_ii] <- NULL
CDE_scRNA$n_op1 <- NULL
CDE_scRNA$CANARY <- NULL
CDE_scRNA$Protocol_.me <- NULL
CDE_scRNA$Surg_Path <- NULL
CDE_scRNA$LDKA <- NULL
write.csv(CDE_scRNA, "/Users/senosam/Documents/Massion_lab/sc_RNA_Seq/DI_CDE_2021JAN13_SA_MF.csv", row.names = FALSE)
```



